Inferring generation-interval distributions from contact tracing data: \\ \emph{in prep, PRSB}

### 2019-04-30 doc/network.tex
small changes

\jd{This is confusing to me. Wouldn't the natural next step be a random network? Homogeneous won't always be shorter than egocentric, I guess. Or are you simulating for a long time? Would it be better to simulate for a shorter time, and add another layer? I was thinking homogeneous $\to$ random $\to$ structured? Doing homogeneous during the exponential period should give us the same as egocentric, right? Maybe we need two figures? Or maybe we don't need a star network at all: first panel of current figure can be done on any network, and second panel can be done on a homogeneous network with a shortish time window?} 
\begin{figure}
\includegraphics[width=\textwidth]{../fig/local_effect.pdf}
\caption{
\textbf{Spatial effects on realized generation intervals.}
Theoretical distributions and means are shown in color (and are the same in each panel, for reference). Simulated distributions and means are shown in black.
(Left) the intrinsic generation-interval distribution corresponds to all contacts by a focal individual, regardless of whether the individual contacted is susceptible (simulated on a star network).
(Middle) the egocentric generation-interval distribution corresponds to the distribution of all infectious contacts by the focal individual with susceptible individuals, in the case where the focal individual is the only possible infector (simulated on a star network).
(Right) realized generation-interval distributions are shorter than egocentric distributions in general, because contacts can be wasted when susceptibles become infected through other routes (simulated on a homogeneous network).
All graphs were generated using 5000 stochastic simulations were on a network with 5 nodes (1 infector and 4 susceptibles).
}
\label{fig:local}
\end{figure}

@@

\swp{Shouldn't we put this earlier?}
In general, spatial reduction in the mean generation interval can be viewed as an effect of susceptible depletion and can be further classified into three levels: egocentric, local, and global.
Egocentric depletion, as discussed previously, is caused by an infected individual making multiple contacts to the same individual.
Local depletion refers to a depletion of susceptible individuals in a household or neighborhood;
this effect can be observed early in an epidemic, especially in a highly structured population, even if most of the population remains susceptible.
Finally, global depletion refers to overall depeletion of susceptibility as a population-level and explains the reduction in realized compared to intrinsic generation intervals that occurs even in a well-mixed population (\fref{censor}).

@@

The intrinsic generation-interval distribution, which describes the expected time distribution of secondary cases, provides a direct link between speed (exponential growth rate, $r$) and strength (reproductive number, $\RR$) of an epidemic \citep{wallinga2007generation, svensson2007note, svensson2015influence, park2019practical}.
However, observed generation-interval distributions can vary depending on how and when it is measured \citep{nishiura2010time, tomba2010some, champredon2015intrinsic, britton2019estimation};
determining which distribution correctly links $r$ and $\RR$ can be challenging.
Here, we explore how observed generation intervals, measured through contact tracing, vary across time and space.
Whereas temporal variations represent observation bias, spatial variations reflect the contact structure of a realized outbreak.
Therefore, correcting for temporal variation, but not for spatial variation, in an observed generation-interval distribution provides the correct link between $r$ and $\RR$, allowing us to estimate the reproductive number $\RR$ that is consistent with the realized infection process.

>>>>

\jd{This is confusing to me. Wouldn't the natural next step be a random network? Homogeneous won't always be shorter than egocentric, I guess. Or are you simulating for a long time? Would it be better to simulate for a shorter time, and add another layer? I was thinking homogeneous $\to$ random $\to$ structured? Doing homogeneous during the exponential period should give us the same as egocentric, right? Maybe we need two figures? Or maybe we don't need a star network at all: first panel of current figure can be done on any network, and second panel can be done on a homogeneous network with a shortish time window?} 
\begin{figure}
\includegraphics[width=\textwidth]{../fig/local_effect.pdf}
\caption{
\textbf{Spatial effects on realized generation intervals.}
Theoretical distributions and means are shown in color (and are the same in each panel, for reference). Simulated distributions and means are shown in black.
(Left) the intrinsic generation-interval distribution corresponds to all contacts by a focal individual, regardless of whether the individual contacted is susceptible (simulated on a star network).
(Middle) the egocentric generation-interval distribution corresponds to the distribution of all infectious contacts by the focal individual with susceptible individuals, in the case where the focal individual is the only possible infector (simulated on a star network).
(Right) realized generation-interval distributions are shorter than egocentric distributions in general, because contacts can be wasted when susceptibles become infected through other routes (simulated on a homogeneous network).
All figures were generated using 5000 stochastic simulations on a network with 5 nodes (1 infector and 4 susceptibles).
}
\label{fig:local}
\end{figure}

@@

In general, spatial reduction in the mean generation interval can be viewed as an effect of susceptible depletion and can be further classified into three levels: egocentric, local, and global.
Egocentric depletion, as discussed previously, is caused by an infected individual making multiple contacts to the same individual.
Local depletion refers to a depletion of susceptible individuals in a household or neighborhood;
this effect can be observed early in an epidemic, especially in a highly structured population, even if most of the population remains susceptible.
Finally, global depletion refers to overall depeletion of susceptibility as a population-level and explains the reduction in realized compared to intrinsic generation intervals that occurs even in a well-mixed population (\fref{censor}).

@@

The intrinsic generation-interval distribution, which describes the expected time distribution of secondary cases, provides a direct link between speed (exponential growth rate, $r$) and strength (reproductive number, $\RR$) of an epidemic \citep{wallinga2007generation, svensson2007note, svensson2015influence, park2019practical}.
However, observed generation-interval distributions can vary depending on how and when it is measured \citep{nishiura2010time, tomba2010some, champredon2015intrinsic, britton2019estimation};
determining which distribution correctly links $r$ and $\RR$ can be challenging.
Here, we explore how observed generation intervals, measured through contact tracing, vary across time and space.
Whereas temporal variations represent observation bias, spatial variations reflect the contact structure of a realized outbreak.
Therefore, correcting for temporal variation, but not for spatial variation, in an observed generation-interval distribution provides a link between $r$ and $\RR$ which allows us to estimate the reproductive number $\RR$ that is consistent with the realized infection process.
